### 2017-10-16 manuscript/manuscript-Word.Rmd
manuscripts - initial commit
manuscript-HTML will produce an HTML preview (easy for editing)

manuscript-Word will produce a word document (to export in a format that co-authors perfer)

Eventually, a latex version will be used, but I won’t add it here, yet.

---
title: "Manuscript_Word"
author: "Januar Harianto"
date: "13/10/2017"
output: 
  word_document:
    reference_docx: "styles/styles-manuscript-Word.docx"
---

```{r setup, include=FALSE}
library(respR)
library(knitr)
library(readr)

# set pander table-layout options
library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

knitr::opts_chunk$set(echo = TRUE)
```